---
title: "BRT_variable_selection_and_correlation"
output: html_document
---
#  Boosted Regression Tree (BRTs) - Variable selection method
#  This R script contains the variable selection method outlined in Irving et al 2019
#  to choose both customised and uniform predictor sets, as well as correlation analysis.
#  the script is an example on a community of 3 species and intended as a guide to the variable selection process
#  

#  please install and upload the following packages

```{r}
install.packages("dismo")
install.packages("rgdal")
install.packages("raster")
install.packages("maptools")
install.packages("reshape")
install.packages("caret")

library(dismo)
library(rgdal)
library(raster)
library(maptools)
library(reshape)
library(caret)
```

#  1st round of BRTs is every predictor set separately e.g. climate only. 
#  script below for one predictor category as an example
#  environmental predictors need to be in raster format (e.g. earthenv.org or IHA metrics are readily downloadable in this format)

#  upload rasters

```{r}
setwd("yourdirectory/rasters")

r1 <- raster("Climate_var1")
r2 <- raster("Climate_var2")
r3 <- raster("Climate_var3")

#  stack all raster
preds <- stack(r1,r2,r3)
```

#  species data needs to be a spatial points dataframe with presence/absence or presence only
#  if presence only, zeros need to be added. 
#  suggestion: add zeros for the remaining sample sites without presences
#  example

```{r}
study_area #  object with all sample site coordinates
presences # object with all presence sites coordinates
#  add a coord code to both dataframes
study_area$coord_code <- paste(study_area$X, "_", study_area$Y, sep="")
presences$coord_code <- paste(presences$X, "_", presences$Y, sep="")

presences$species <- 1 # add column of presences

## merge both dataframes by coord code
data <- merge(study_area, presences, by="coord_code", all=T)
#  remove unwanted columns - coordinates and species needed only
data <- data[, -c(3,4,5,6)] 
data[is.na(data)] <- 0 ## change NAs to 0. dataframe will have binary info now

#  convert to spatial points dataframe

coordinates(data) <- C("X", "Y")
#  extract environmental data from rasters at sampling sites
sx2 <-raster::extract(preds,data) ## extract info from environamtal raster at sampling sites
df <- cbind(data, sx2) ## combine data for species and variables, creates format dataframe as needed for brt
df <- na.omit(df)

df # dataframe formatted for BRTs
```
#  dataframe will now be formatted in species presense/absence (1/0), with coordinates 
# of sampling sites and environmental variables. Example: 

# # X        Y                    species    Climate_var1    Climate_var2    Climate_var3
# 1 10.00417 50.89583                   1              84              81              32
# 2 10.00417 51.21250                   0              87              80              32
# 3 10.00417 51.25417                   0              84              82              33
# 4 10.00417 52.43750                   1              74              70              29
# 5 10.01250 50.62917                   0              68              67              29
# 6 10.01250 50.90417                   0              78              71              30

#  BRTs

```{r}
#  creat dataframe for BRT coefficients
coefs <- data.frame(matrix(ncol=6))
colnames(coefs) <- c("taxa", "no_of_trees", "dev_mean", "cor_mean", "discrim_mean", "cv_thres")
#  more coefficients are available check cv.statistics
tbrt <- gbm.step(data, gbm.x = 4:6, gbm.y = 3, family ="bernoulli", tree.complexity = 5, learning.rate = 0.0001, bag.fraction = 0.5) 

#  df = dataframe as above
#  gbm.x = 4:6 - the environmental variables in the dataframe
#  gbm.y = 3 - the species presence/absence
#  BRT criteria can be changed according to needs. Please check ?gbm.step and Elith et al (2008) 

var_imp <- as.data.frame(summary(tbrt)) # Variable importance dataframe
#  var_imp should be formatted as below

# var           rel.inf
# Climate_var1  25.2409052
# Climate_var3  12.9871157
# Climate_var2  8.9114047

tbrt$cv.statistics$cv.threshold #  BRT coefficients

#  puts coefficients information in a dataframe
coefs[1,1] <- "species name" #  add species name
coefs[1,2] <- tbrt$n.trees
coefs[1,3] <- tbrt$cv.statistics$deviance.mean
coefs[1,4] <- tbrt$cv.statistics$correlation.mean
coefs[1,5] <- tbrt$cv.statistics$discrimination.mean
coefs[1,6] <- tbrt$cv.statistics$cv.threshold

#  save both 

save(var_imp, file="yourdirectory/filename_var_imp.RData")
save(coefs, file="yourdirectory/filename2_coefs.RData")


```

#   repeat for each species in the community
#  below is an example for each loop for a guide

```{r}
species # list of species names as in filenames
study_area #  dataframe with all sample site coordinates
study_area$coord_code <- paste(study_area$X, "_", study_area$Y, sep="")#  add a coord code

for( i in 1: length(species)) {
  
  
  # if presence only data
  presences <- read.csv("yourdirectory/file_species1.csv") # upload file of species presences
  
  #  add a coord code 
  
  presences$coord_code <- paste(presences$X, "_", presences$Y, sep="")
  
  presences$species <- 1 # add column of presences
  
  ## merge both dataframes by coord code
  data <- merge(study_area, presences, by="coord_code", all=T)
  #  remove unwanted columns - coordinates and species needed only
  data <- data[, -c(3,4,5,6)] 
  data[is.na(data)] <- 0 ## change NAs to 0. dataframe will have binary info now
  
  # if presence/absence data 
  data  <- read.csv("yourdirectory/file_species1.csv") 
  #  convert to spatial points dataframe
  
  coordinates(data) <- C("X", "Y")
  #  extract environmental data from rasters at sampling sites
  sx2 <-raster::extract(preds,data) ## extract info from environamtal raster at sampling sites
  df <- cbind(data, sx2) ## combine data for species and variables, creates format dataframe as needed for brt
  df <- na.omit(df)
  
  
  tbrt <- gbm.step(df, gbm.x = 4:6, gbm.y = 3, family ="bernoulli", tree.complexity = 5, learning.rate = 0.0001, bag.fraction = 0.5) 
  
  var_imp <- as.data.frame(summary(tbrt)) # Variable importance dataframe
  
  tbrt$cv.statistics$cv.threshold #  BRT coefficients
  
  #  puts coefficients information in a dataframe
  coefs[1,1] <- "species name" #  add species name
  coefs[1,2] <- tbrt$n.trees
  coefs[1,3] <- tbrt$cv.statistics$deviance.mean
  coefs[1,4] <- tbrt$cv.statistics$correlation.mean
  coefs[1,5] <- tbrt$cv.statistics$discrimination.mean
  coefs[1,6] <- tbrt$cv.statistics$cv.threshold
  
  #  save both 
  
  save(var_imp2, file=paste("yourdirectory/filename_var_imp2_", species[i],".RData", sep=""))
  save(coefs2, file=paste("yourdirectory/filename_coefs2_", species[i],".RData", sep=""))
  
  rm(var_imp)
  rm(coefs)
  
  
}


#   repeat for each predictor category

```

#   repeat for each predictor category

```{r}
#  combine variable importance for each species

setwd("yourdirectory")


cl <- list.files(pattern="var_imp") #  list all species files. change pattern="var_imp" according to filename

load(file=paste(cl[1])) ## var_imp (name of file)

var_impx <- var_imp ## new var_imp object to accumulate dataframes

var_impx <- var_impx[order(var_impx$var),] ## order by variable name 

for(c in 2:length(cl)) {
  
  cat("Running site", c, "\n")
  load(file=paste(cl[c])) ## var_imp ## load file
  colnames(var_imp)[1] <- "var" ## 1st column called var 
  var_imp<- var_imp[order(var_imp$var),]## order by variable name
  
  
  sp <- colnames(var_imp)[2] ## get species name
  
  var_imp <- var_imp[, -1] ## remove variable column
  
  var_impx <- cbind(var_impx, var_imp)  ## combine by column
  colnames(var_impx)[1+c] <- sp ## add species name
  
  
}

#  save file
save(var_impx, file="yourdirectory/filename.RData")

#           var               species 1        species 2       species 3
# 1 climate var1            3.196427          9.731135         6.516256
# 2 climate var2            5.743172          1.963038         4.256472
# 3 climate var3           2.299511          1.202297         2.243538

```
#  repeat for all predictor categories

#  for customised predictor set the variable importance values for each species will be taken
#  to create a uniform predictor set 

```{r}
var_impx$mean <- rowMeans(var_impx[2:4]) ## average across all variables

```

#           var               species 1        species 2       species 3
# 1 climate var1            3.196427          9.731135         6.516256
# 2 climate var2            5.743172          1.963038         4.256472
# 3 climate var3           2.299511          1.202297         2.243538#    
# mean
 # 6.481273
 # 3.987561
 # 1.915116

#  mean column is the variable importance for the community

# subset only community variable importance and take most important variables

```{r}
sp1 <- clim2[, c(1, 5)]
sp1 <- sp1[order(sp1[,2]),]
#  extract the most important 30 % variables 
#  in this example it is 1, however if the predictor set has e.g. 19 variables, the 
#  top 30% would be 6
topvars <- tail(sp1, 1) ## 30% of 3

#  topvars contains the most important variables for the community for this predictor category


```
```{r}
#  save file
save(topvars, file="yourdirectory/filename.RData")
```
#  2nd run of BRTs - all remaining predictors i.e. most important 30% from each predictor category

# # X        Y                    species    Climate_var1    hydrology_var1    topography_var1
# 1 10.00417 50.89583                   1              84              81              32
# 2 10.00417 51.21250                   0              87              80              32
# 3 10.00417 51.25417                   0              84              82              33
# 4 10.00417 52.43750                   1              74              70              29
# 5 10.01250 50.62917                   0              68              67              29
# 6 10.01250 50.90417                   0              78              71              30

```{r}
#  upload rasters as above but with all top predictors
#  get list of all predictors

setwd("youdirectory/climaterasters")
cp <- list.files(pattern=".tif")


setwd("youdirectory/hydrologyrasters")
hp <- list.files(pattern=".tif")
hp

setwd("youdirectory/topographyrasters")
tp <- list.files(pattern=".tif")

#  stack rasters from each category

load("yourdirectory/filename.RData") ## topvars climate
canx <- topvars
var_list <- droplevels(canx$var)


setwd("youdirectory/climaterasters")

c <- grep(paste0(var_list, ".tif", collapse="|"),   cp)
cpx <- cp[c]
cpr <- stack(cpx)


load("yourdirectory/filename.RData") ## topvars hydrology
hanx <- topvars
var_list <- droplevels(hanx$var)


setwd("youdirectory/hydrologyrasters")

h <- grep(paste0(var_list, ".tif", collapse="|"),   hp)
hpx <- hp[h]
hpr <- stack(hpx)

load("yourdirectory/filename.RData") ## topvars topography
tanx <- topvars
var_list <- droplevels(tanx$var)
var_list

setwd("youdirectory/topographyrasters")

t <- grep(paste0(var_list, ".tif", collapse="|"),   tp)
tpx <- tp[t]
tpr <- stack(tpx)

preds <- stack(cpr,hpr.tpr)


```

# species data as above (1st run BRTs)

```{r}
#  for loop to run each species

species # list of species names as in filenames
study_area #  dataframe with all sample site coordinates
study_area$coord_code <- paste(study_area$X, "_", study_area$Y, sep="")#  add a coord code

for( i in 1: length(species)) {
  
  
  # if presence only data
  presences <- read.csv("yourdirectory/file_species1.csv") # upload file of species presences
  
  #  add a coord code 
  
  presences$coord_code <- paste(presences$X, "_", presences$Y, sep="")
  
  presences$species <- 1 # add column of presences
  
  ## merge both dataframes by coord code
  data <- merge(study_area, presences, by="coord_code", all=T)
  #  remove unwanted columns - coordinates and species needed only
  data <- data[, -c(3,4,5,6)] 
  data[is.na(data)] <- 0 ## change NAs to 0. dataframe will have binary info now
  
  # if presence/absence data 
  data  <- read.csv("yourdirectory/file_species1.csv") 
  #  convert to spatial points dataframe
  
  coordinates(data) <- C("X", "Y")
  #  extract environmental data from rasters at sampling sites
  sx2 <-raster::extract(preds,data) ## extract info from environamtal raster at sampling sites
  df <- cbind(data, sx2) ## combine data for species and variables, creates format dataframe as needed for brt
  df <- na.omit(df)
  
  
  tbrt <- gbm.step(df, gbm.x = 4:6, gbm.y = 3, family ="bernoulli", tree.complexity = 5, learning.rate = 0.0001, bag.fraction = 0.5) 
  
  var_imp <- as.data.frame(summary(tbrt)) # Variable importance dataframe
  
  tbrt$cv.statistics$cv.threshold #  BRT coefficients
  
  #  puts coefficients information in a dataframe
  coefs[1,1] <- "species name" #  add species name
  coefs[1,2] <- tbrt$n.trees
  coefs[1,3] <- tbrt$cv.statistics$deviance.mean
  coefs[1,4] <- tbrt$cv.statistics$correlation.mean
  coefs[1,5] <- tbrt$cv.statistics$discrimination.mean
  coefs[1,6] <- tbrt$cv.statistics$cv.threshold
  
  #  save both 
  
  save(var_imp2, file=paste("yourdirectory/filename_var_imp2_", species[i],".RData", sep=""))
  save(coefs2, file=paste("yourdirectory/filename_coefs2_", species[i],".RData", sep=""))
  
  rm(var_imp)
  rm(coefs)
  
  
}

```



#  merge all species variable importance together in one file for correlation

```{r}
#  directory where var_imp from 2nd brt run saved
setwd("yourdirectory/file")

ul <- list.files(pattern="var_imp")# list of all species variable importance files

```
```{r}
## make dataframe to add all dfs together
b_resx <- data.frame(matrix(nrow=3, ncol=2)) # nrow is number of variables
colnames(b_resx)[1] <- "var"

```

```{r}
### format first file

setwd("youdirectory/file")
load(file=paste(ul[1])) ## var_imp
#  all variables as column 
b_resx[, 1] <-  var_imp$var 

b_resx <- merge(b_resx, var_imp, by="var", all=T)# merge dataframes
b_resx <- b_resx[, c(1, 3)] # take only needed columns
sp <- colnames(var_imp)[2] #  species name
colnames(b_resx)[2] <- sp # add species name to new dataframe

# loop all files together

for(c in 2:length(ul)) {
  
  cat("Running site", c, "\n")
  load(file=paste(ul[c])) ## var_imp ## load file
  
  colnames(var_imp)[1] <- "var" ## 1st colum called var - taxa with no brt have no name
  var_imp <- var_imp[order(var_imp$var),]## order by variable name
  # dim(b_res)
  bx <- dim(var_imp)[1]
  
  sp <- colnames(var_imp)[2] ## get species name
  
    var_imp <- var_imp[, -1] ## remove variable column
  
  
    b_resx <- cbind(b_resx, var_imp)  ## combine by column
  
    colnames(b_resx)[1+c] <- sp ## add species name
  
  
}


write.csv(b_resx, "yourdirectory/all_var_imp_uniform.csv")

```

#  correlation analysis for uniform predictor set
# for this predictor set the correlation only need to be done once

```{r}
#  upload variable importance for all species
all_preds <- read.csv("yourdirectory/all_var_imp_uniform.csv")
all_preds <- all_preds[, -1] #  remove x column

all_preds$mean <- rowMeans(all_preds[2:4])  #  average over species for community variable importance


## make df for correlation
vdf <- data.frame(matrix(nrow=3, ncol=1))
vdf[,1] <- all_preds[, 1] # add variables as first column

colnames(vdf)[1] <- "Variables"
 
top_vars <- all_preds[, c(1, 5)] # community variable importance


var_list <- all_preds[, 1] # list all variables

```

#  get list of all predictors
#  this was done above, so may not be needed again here

```{r}


setwd("youdirectory/climaterasters")
cp <- list.files(pattern=".tif")


setwd("youdirectory/hydrologyrasters")
hp <- list.files(pattern=".tif")


setwd("youdirectory/topographyrasters")
tp <- list.files(pattern=".tif")

#  load variables according to var_list


setwd("youdirectory/climaterasters")

c <- grep(paste0(var_list, ".tif", collapse="|"),   cp)
cpx <- cp[c]
cpr <- stack(cpx)


setwd("youdirectory/hydrologyrasters")

h <- grep(paste0(var_list, ".tif", collapse="|"),   hp)
hpx <- hp[h]
hpr <- stack(hpx)


setwd("youdirectory/topographyrasters")

t <- grep(paste0(var_list, ".tif", collapse="|"),   tp)
tpx <- tp[t]
tpr <- stack(tpx)

#  stack all predictors
preds <- stack(cpr,hpr.tpr)

#  covert to dataframe
preds_df <- as.data.frame(preds)
preds_df <- na.omit(preds_df)
```

###### multicolinearality

```{r}
df2 = cor(preds_df)

df2 <- as.data.frame(df2)


df2$var <- rownames(df2)
cor_imp <- merge(topvars, df2, by="var")

## make dataframe in order of relative importance
cor_imp <- cor_imp[order(cor_imp[,2], decreasing=TRUE),]

## remove minus signs from data frame 
cor_imp[] <- lapply(cor_imp, function(x) gsub("-", "", x ))
## change from character to number
cor_imp[, 2:25] <- lapply(cor_imp[,2:25], function(x) as.numeric(as.character(x)))

### if over 0.7 take the variable with the higher importance

## reshape dataframe
colnames(cor_imp)[2] <- "importance"
cor_imp <- melt(cor_imp, id=c("var", "importance"))
cor_imp$importance2 <- NA

v <- cor_imp$var
## df column for importance of variable 2 
for(d in 1: length(v)) {
  
  v1 <- v[d]
  imp <- cor_imp$importance[d] 
  v2 <- cor_imp$variable %in% v1
  cor_imp$importance2[v2] <- paste(imp)
  
}


cor_imp <- na.omit(cor_imp)

vals <- cor_imp$value

### find correlations over 0.7

for (c in 1: length(vals)) {
  
  cv <- vals[c]
  if ( cv > 0.7 && cv < 1){
    cor_imp$corr[c] <- paste("true")
  } else {
    cor_imp$corr[c] <- paste("false")
  }
}


## subset df to only correlations over 0.7
varcors <- subset(cor_imp, corr=="true")

left_vars <- varcors$var

### take the variable of higher relative importance - make list of losing predictors then remove from list of variables.

for (l in 1: length(left_vars)) {
  
  ### first separate pairwise correlations
  ## variable 1
  lvx <- left_vars[l]
  
  ## variable 2
  ov <- varcors$variable[l]

  ## variable 2 in list to find var importance value
  ovx <- match(ov, left_vars)

  if (varcors[l, 2] < varcors[ovx, 2]) {
    
    varcors$remove[l] <- paste(lvx)
    
  } else { 
    
    varcors$remove[l] <- paste(ov)
    
  }
  
}

## remove any variables that are correlated
vb <- unique(varcors$remove)
sd <- vars %in% vb
del <- vars[!sd]
del ## remaining variables to use - subset from preds dataframe

resx <- subset(cor_imp, var %in% del)
resx <- resx[, 1:2]
resx <- resx[!duplicated(resx),]

write.csv(resx, file="youdirectory/all_preds_for_sdms_uniform.csv")

```

Uniform predictor set now ready for input into SDM 

#  customised predictor set. 
#  this is a little more complex as different predictors are taken for each species

```{r}


#  get list of all predictors

setwd("youdirectory/climaterasters")
cp <- list.files(pattern=".tif")


setwd("youdirectory/hydrologyrasters")
hp <- list.files(pattern=".tif")


setwd("youdirectory/topographyrasters")
tp <- list.files(pattern=".tif")

```

loop around all species

```{r}

species # list of species names as in filenames
study_area #  dataframe with all sample site coordinates
study_area$coord_code <- paste(study_area$X, "_", study_area$Y, sep="")#  add a coord code

for( i in 1: length(species)) {
  
  
  # if presence only data
  presences <- read.csv("yourdirectory/file_species1.csv") # upload file of species presences
  
  #  add a coord code 
  
  presences$coord_code <- paste(presences$X, "_", presences$Y, sep="")
  
  presences$species <- 1 # add column of presences
  
  ## merge both dataframes by coord code
  data <- merge(study_area, presences, by="coord_code", all=T)
  #  remove unwanted columns - coordinates and species needed only
  data <- data[, -c(3,4,5,6)] 
  data[is.na(data)] <- 0 ## change NAs to 0. dataframe will have binary info now
  
  # if presence/absence data 
  data  <- read.csv("yourdirectory/file_species1.csv") 
  #  convert to spatial points dataframe
  
  #  upload predictor data
  
  #  stack rasters from each category
  
  load(paste("yourdirectory/climate_filename", species[i], ".RData", sep="")) ## topvars climate for species i
  canx <- topvars
  var_list <- droplevels(canx$var)
  var_list
  
  setwd("youdirectory/climaterasters")
  
  c <- grep(paste0(var_list, ".tif", collapse="|"),   cp)
  cpx <- cp[c]
  cpr <- stack(cpx)
  
  
  load(paste("yourdirectory/hydrology_filename", species[i], ".RData", sep="")) ## topvars hydrology for species i
  hanx <- topvars
  var_list <- droplevels(hanx$var)
  var_list
  
  setwd("youdirectory/hydrologyrasters")
  
  h <- grep(paste0(var_list, ".tif", collapse="|"),   hp)
  hpx <- hp[h]
  hpr <- stack(hpx)
  
  load(paste("yourdirectory/topography_filename", species[i], ".RData", sep="")) ## topvars topography for species i
  tanx <- topvars
  var_list <- droplevels(tanx$var)
  var_list
  
  setwd("youdirectory/topographyrasters")
  
  t <- grep(paste0(var_list, ".tif", collapse="|"),   tp)
  tpx <- tp[t]
  tpr <- stack(tpx)
  
  preds <- stack(cpr,hpr.tpr)
  
  coordinates(data) <- C("X", "Y")
  #  extract environmental data from rasters at sampling sites
  sx2 <-raster::extract(preds,data) ## extract info from environamtal raster at sampling sites
  df <- cbind(data, sx2) ## combine data for species and variables, creates format dataframe as needed for brt
  df <- na.omit(df)
  
  
  tbrt <- gbm.step(df, gbm.x = 4:6, gbm.y = 3, family ="bernoulli", tree.complexity = 5, learning.rate = 0.0001, bag.fraction = 0.5) 
  
  var_imp <- as.data.frame(summary(tbrt)) # Variable importance dataframe
  
  tbrt$cv.statistics$cv.threshold #  BRT coefficients
  
  #  puts coefficients information in a dataframe
  coefs[1,1] <- "species name" #  add species name
  coefs[1,2] <- tbrt$n.trees
  coefs[1,3] <- tbrt$cv.statistics$deviance.mean
  coefs[1,4] <- tbrt$cv.statistics$correlation.mean
  coefs[1,5] <- tbrt$cv.statistics$discrimination.mean
  coefs[1,6] <- tbrt$cv.statistics$cv.threshold
  
  #  save both 
  
  save(var_imp2, file=paste("yourdirectory/filename_var_imp2_", species[i],".RData", sep=""))
  save(coefs2, file=paste("yourdirectory/filename_coefs2_", species[i],".RData", sep=""))
  
  rm(var_imp)
  rm(coefs)
  
  
}


```

#  custom predictor set - merge all variable importance dataframes

```{r}
rels <- c("Climate_var1",    "hydrology_var1",  "topography_var1") # list all variables

## make dataframe to add all dfs together
b_resx <- data.frame(matrix(nrow=3, ncol=2))
b_resx[, 1] <-  rels 
colnames(b_resx)[1] <- "var"

### format first file

setwd("yourdirectory/file")

#  where custom variable importance from 2nd brt run is saved
cl <- list.files(pattern="var_imp") # list species files

load(file=paste(cl[1])) ## b_res
head(var_imp)
length(b_res$var) ## 29

b_resx <- merge(b_resx, var_imp, by="var", all=T)# merge dataframes
b_resx <- b_resx[, c(1, 3)] # take only needed columns
sp <- colnames(var_imp)[2] #  species name
colnames(b_resx)[2] <- sp # add species name to new dataframe

for(c in 2:length(cl)) {
  
  cat("Running site", c, "\n")
  load(file=paste(cl[c])) ## var_imp ## load file
  
  colnames(var_imp)[1] <- "var" 
  b_res <- var_imp[order(var_imp$var),]## order by variable name
  
  bx <- dim(var_imp)[1]
  
  sp <- colnames(var_imp)[2] ## get species name
  
    b_resx <- merge(b_resx, var_imp, by="var", all=T) ## combine by column
  
  colnames(b_resx)[1+c] <- sp ## add species name
  
}

write.csv(b_resx, "yourdirectory/all_var_imp_custom.csv")


```

#  customised predictor set correlation analysis, one analysis for each species

```{r}

### variable importance data  
setwd("youdirectory/")

all_preds <- read.csv("yourdirectory/all_var_imp_custom.csv")
all_preds <- all_preds[, -1]

## make df for correlation
vdf <- data.frame(matrix(nrow=3, ncol=1))
vdf[,1] <- all_preds[, 1] # add variables as first column

colnames(vdf)[1] <- "Variables"

```

Correlation for all species 

```{r}
species # list of species

for(t in 1:length(species)) {
  
  
  topvars <- all_preds[, c(1, 1+t)]
  
  topvars <- na.omit(topvars)
  
  var_list <- all_preds[, 1] # list all variables
  

  ### extract variable data
  setwd("youdirectory/climaterasters")
  
  c <- grep(paste0(var_list, ".tif", collapse="|"),   cp)
  cpx <- cp[c]
  cpr <- stack(cpx)
  
  
  setwd("youdirectory/hydrologyrasters")
  
  h <- grep(paste0(var_list, ".tif", collapse="|"),   hp)
  hpx <- hp[h]
  hpr <- stack(hpx)
  
  
  setwd("youdirectory/topographyrasters")
  
  t <- grep(paste0(var_list, ".tif", collapse="|"),   tp)
  tpx <- tp[t]
  tpr <- stack(tpx)
  
  #  stack all predictors
  preds <- stack(cpr,hpr.tpr)
  
  #  predi tors as data frame
  preds_df <- as.data.frame(preds)
  preds_df <- na.omit(preds_df)
  
  df2 = cor(preds_df)
  
  df2 <- as.data.frame(df2)
  
  
  df2$var <- rownames(df2)
  cor_imp <- merge(topvars, df2, by="var")
  
  ## make dataframe in order of relative importance
  cor_imp <- cor_imp[order(cor_imp[,2], decreasing=TRUE),]
  
  ## remove minus signs from data frame 
  cor_imp[] <- lapply(cor_imp, function(x) gsub("-", "", x ))
  ## change from character to number
  cor_imp[, 2:4] <- lapply(cor_imp[,2:4], function(x) as.numeric(as.character(x)))
  
  #  indexing in cor_imp[, 2:4] will need to be adjusted according to your dataframe dimensions
  ### if over 0.7 take the variable with the higher importance
  
  ## reshape dataframe
  
  colnames(cor_imp)[2] <- "importance"
  cor_imp <- melt(cor_imp, id=c("var", "importance"))
  cor_imp$importance2 <- NA
  
  v <- cor_imp$var
  ## df column for importance of variable 2 
  for(d in 1: length(v)) {
    
    v1 <- v[d]
    imp <- cor_imp$importance[d] 
    v2 <- cor_imp$variable %in% v1
    cor_imp$importance2[v2] <- paste(imp)
    
  }
  
  
  cor_imp <- na.omit(cor_imp)
  
  vals <- cor_imp$value
  
  ### find correlations over 0.7
  
  for (c in 1: length(vals)) {
    
    cv <- vals[c]
    if ( cv > 0.7 && cv < 1){
      cor_imp$corr[c] <- paste("true")
    } else {
      cor_imp$corr[c] <- paste("false")
    }
  }
  
  
  ## subset df to only correlations over 0.7
  varcors <- subset(cor_imp, corr=="true")
  
  left_vars <- varcors$var
  
  ### take the variable of higher relative importance - make list of losing predictors then remove from list of variables.
  
  for (l in 1: length(left_vars)) {
    
    ### first separate pairwise correlations
    ## variable 1
    lvx <- left_vars[l]
    
    ## variable 2
    ov <- varcors$variable[l]
    
    ## variable 2 in list to find var importance value
    ovx <- match(ov, left_vars)
    
    if (varcors[l, 2] < varcors[ovx, 2]) {
      
      varcors$remove[l] <- paste(lvx)
      
    } else { 
      
      varcors$remove[l] <- paste(ov)
      
    }
    
  }
  
  ## remove any variables that are correlated
  vb <- unique(varcors$remove)
  sd <- vars %in% vb
  del <- vars[!sd]
  del ## remaining variables to use - subset from preds dataframe
  
  
  ## subset from cor df and remove duplicates
  resx <- subset(cor_imp, var %in% del)
  resx <- resx[, 1:2]
  resx <- resx[!duplicated(resx),]
  
  
  ## merge with all species variable importance
  colnames(resx)[2] <- paste(colnames(sp1)[2])
  vdf <- merge(vdf, resx, by.x = "Variables", by.y = "var", all =T)
  
} 
  

write.csv(vdf, file="youdirectory/preds_for_sdms_custom.csv")

```

Custom predictor set now ready for input into SDM 